Linear analysis of the vertical shear instability in astrophysical disks

### 2015-02-23 appendix.tex
new subsection growth time vs rad for mmsn

However, in the presence of vertical shear $q\neq0$,
Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
$\khat$, which implies the possibility of instability, regardless of
how much $\gamma$ deviates from unity. This is contradicts the second 
Solberg-Hoiland criterion (Eq. \ref{solberg2}), which states that
instability in the adiabatic limit requires the disk to be close to
neutrally stratified. %also need real freq^2 in adia limit 
This is because the VGSB retains the destabilizing effect of vertical shear
($\p_z\Omega$) from the global disk, but ignores the global radial disk
structure responsible for it. Nevertheless, we demonstrate below that
this inconsistency is unimportant for the vertical shear instability,
which occurs for $\beta\ll1$.

@@

\section{Axisymmetric instability in vertically 
  isothermal disks in global cylindrical geometry}\label{global_vertiso}
Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
all times, so in the perturbed state $\delta P =
c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
The axisymmetric linearized fluid equations in cylindrical
geometry is then 
\begin{align}
 \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
 \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
 \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
   z}\left(\frac{\delta\rho}{\rho}\right),\\
 \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right),
\end{align}
and we recall $j=r^2\Omega$. We can eliminate $\delta v_\phi$ from these equations to obtain
\begin{align}
  -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
  \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
  \delta v_r^*\delta v_z =
  c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
  =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral} 
\end{align}
At this point if we make the anelastic approximation by assuming
$\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
equation \citep[as argued by][]{nelson13}, 
then the only contribution to instability comes from the third term on
the left hand side, i.e. the explicit vertical shear term
$\p_zj$. This is the same conclusion reached in
\S\ref{integral_relation}.

The term proportional to $\kappa^2$ is stabilizing. Because radial shear
is much larger than vertical shear, this implies that significant
instability can only be achieved with large vertical
motions (i.e. $|\delta v_z|\gg|\delta v_r|$).

If we integrate Eq. \ref{global_integral} over the fluid volume 
and assume boundary terms vanish when integrating by parts, we obtain 
\begin{align}
  -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
  \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
  \delta v_r^*\delta v_z dV = -\int
  \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
  dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
  \bm{v}^*\cdot\nabla c_s^2.
\end{align}
Note that the treatment so far is in fact valid for any prescribed
$c_s(r,z)$. However, to proceed further we specialize to the case
$c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
disks. Recalling $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
c_s^2$  for a vertically isothermal disk (Eq. \ref{vertical_shear}), we obtain 
\begin{align}
  \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
  \int \kappa^2\rho|\delta v_r|^2 dV + \int
  \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
  dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
    v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
\end{align}
We see that instability is only possible if
$dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
decrease outwards in astrophysical disks, instability is associated
with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
clear that $\sigma^2$ will be complex in general.

>>>>

However, in the presence of vertical shear $q\neq0$,
Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
$\khat$, which implies the possibility of instability, regardless of
how much $\gamma$ deviates from unity. This is contradicts the second 
Solberg-Hoiland criterion (Eq. \ref{solberg2}), which states that
instability in the adiabatic limit requires the disk to be close to
neutrally stratified. %also need real freq^2 in adia limit 
This is because the VGSB retains the vertical shear
($\p_z\Omega$) from the global disk, but ignores the global radial disk
structure responsible for it. Nevertheless, we demonstrate below that
this inconsistency is unimportant for the vertical shear instability,
which occurs for $\beta\ll1$.

@@

% \section{Axisymmetric instability in vertically 
%   isothermal disks in global cylindrical geometry}\label{global_vertiso}
% Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
% all times, so in the perturbed state $\delta P =
% c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
% The axisymmetric linearized fluid equations in cylindrical
% geometry is then 
% \begin{align}
%  \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
%  \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
%  \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
%    z}\left(\frac{\delta\rho}{\rho}\right),\\
%  \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right),
% \end{align}
% and we recall $j=r^2\Omega$. We can eliminate $\delta v_\phi$ from these equations to obtain
% \begin{align}
%   -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
%   \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
%   \delta v_r^*\delta v_z =
%   c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
%   =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral} 
% \end{align}
% At this point if we make the anelastic approximation by assuming
% $\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
% equation \citep[as argued by][]{nelson13}, 
% then the only contribution to instability comes from the third term on
% the left hand side, i.e. the explicit vertical shear term
% $\p_zj$. This is the same conclusion reached in
% \S\ref{integral_relation}.

% The term proportional to $\kappa^2$ is stabilizing. Because radial shear
% is much larger than vertical shear, this implies that significant
% instability can only be achieved with large vertical
% motions (i.e. $|\delta v_z|\gg|\delta v_r|$).

% If we integrate Eq. \ref{global_integral} over the fluid volume 
% and assume boundary terms vanish when integrating by parts, we obtain 
% \begin{align}
%   -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
%   \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
%   \delta v_r^*\delta v_z dV = -\int
%   \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
%   dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
%   \bm{v}^*\cdot\nabla c_s^2.
% \end{align}
% Note that the treatment so far is in fact valid for any prescribed
% $c_s(r,z)$. However, to proceed further we specialize to the case
% $c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
% disks. Recalling $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
% c_s^2$  for a vertically isothermal disk (Eq. \ref{vertical_shear}), we obtain 
% \begin{align}
%   \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
%   \int \kappa^2\rho|\delta v_r|^2 dV + \int
%   \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
%   dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
%     v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
% \end{align}
% We see that instability is only possible if
% $dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
% decrease outwards in astrophysical disks, instability is associated
% with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
% clear that $\sigma^2$ will be complex in general.

### 2015-02-23 application.tex
new subsection growth time vs rad for mmsn

\subsubsection{Vertically-dependent thermal relaxation} 
Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
we can restore the $z$ dependence in $\beta$ (due to the $\rho$
dependence in  $\eta$, see Eq. \ref{eta_def}) by making the transformation
$\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv 
\tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
$\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example
calculation above with $\beta\to\tilde{\beta}$, but found similar
results. 
%We find growth rates
%were reduced, but at most by $\sim 15\%$ at $\khat=1$. Eigenfunctions
%were similar.

>>>>

\subsection{Applicability of the VSI to the MMSN}
To obtain an overall picture of the relevance of the VSI to the MMSN,
we calculate characteristic growth times $t_\mathrm{grow} = 1/\nu$ of 
the fundamental mode in units of the local Keplerian orbital period
$P_\mathrm{orb}= 2\pi/\Omega_k$. The results are shown in 
Fig. \ref{mmsn_overall} for perturbation wavenumbers $5\lesssim
\khat\lesssim 50$. We consider this range because for  
$\khat=O(1)$ or $\khat=O(10^2)$ we found eigenfunctions have large
amplitudes near the vertical boundaries, so such modes may be affected
by boundary conditions.

In the MMSN the fundamental VSI operates at tens of AU, with
characteristic growth times of tens of orbits. %In our example, the VSI
% is most favored at $r\simeq 20\mathrm{AU}$ with a growth time
% $t_\mathrm{grow}\simeq 30$ orbits.
Fig. \ref{mmsn_overall} indicates that with decreasing radius, the VSI
shifts to smaller scales (larger $\khat$), since the disk becomes optically
thick so only small-scale disturbances have sufficiently short thermal
relaxation timescales. At larger radii ($r\gtrsim 
40\mathrm{AU}$,) the VSI can operate on larger scales (smaller
$\khat$) and growth times do not depend strongly on
$\khat$, since the disk is optically thin. However, approaching
$r\simeq 100\mathrm{AU}$ growth times become long, at $O(100)$ orbits so
the VSI may not be dynamically important in the outermost parts of the
MMSN.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigen_compare_grow.ps}
  \caption{Characteristic growth times of the fundamental VSI mode in
    the MMSN for a range of perturbation wavenumbers.  
    \label{mmsn_overall}}    
\end{figure}

 % \subsection{Vertically-dependent thermal relaxation} 
% Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
% we can restore the $z$ dependence in $\beta$ (due to the $\rho$
% dependence in  $\eta$, see Eq. \ref{eta_def}) by making the transformation
% $\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv 
% \tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
% $\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example
% calculation in \S\ref{mmsn_example} with $\beta\to\tilde{\beta}$, but
% found similar results.

%We find growth rates
%were reduced, but at most by $\sim 15\%$ at $\khat=1$. Eigenfunctions
%were similar.

### 2015-02-23 linear_adia.tex
new subsection growth time vs rad for mmsn

Here we estimate the maximum thermal relaxation timescale
$\beta$ that allows the fundamental VSI for $\khat\gg1$.   
% Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
% fundamental isothermal VSI tends to a constant for large
% $\khat$. Using this clue, we can calculate this critical $\beta$ by
% considering $\khat\to\infty$. We proceed as follows. 
At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
is real. The dispersion relation for the fundamental mode is then
\begin{align}\label{relax_disp_fund}
  0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
\end{align} 
Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} in
the limit $\khat\to\infty$ we find
\begin{align}
  0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
  \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
  &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
  0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
  - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3.\label{relax_cond2}
\end{align}
We seek solutions with
\begin{align}
  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
\end{align}
Then
\begin{align}
  \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
\end{align}
from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
that
\begin{align}
  \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
        q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
\end{align}
at marginal stability for perturbations with large $\khat$. 
For a thin disk $\epsilon\ll 1$, we infer 
\begin{align}\label{iso_vsi_cond}
  \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
  \beta_\mathrm{crit} 
\end{align}
is required for the fundamental VSI to operate. From our discussion
for isothermal perturbations (\S\ref{iso_discuss}), we expect
$\beta<\beta_\mathrm{crit}$ to be a sufficient condition for the
VSI. That is, when $\beta<\beta_\mathrm{crit}$, perturbations are
effectively isothermal. For a thin disk 
$\beta_\mathrm{crit}\ll 1$, which suggests that finite thermal
relaxation rapidly stabilizes the VSI.

>>>>

Here we estimate the maximum thermal relaxation timescale
$\beta$ that allows the fundamental VSI for $\khat\gg1$.   
% Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
% fundamental isothermal VSI tends to a constant for large
% $\khat$. Using this clue, we can calculate this critical $\beta$ by
% considering $\khat\to\infty$. We proceed as follows. 
At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
is real. The dispersion relation for the fundamental mode is then
\begin{align}\label{relax_disp_fund}
  0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
\end{align} 
Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
$\khat\gg1$, we find
\begin{align}
  0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
  \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
  &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
  0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
  - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3.\label{relax_cond2}
\end{align}
This is a pair of simultaneous equations for
$\hat{\omega}$ and $\beta$.%  A simple solution can be obtained by the
% following considerations.
Now, our discussion of stable solutions 
without vertical shear yield a dispersion
relation $\hat{\omega}\propto \khat^{-1}$ for $\khat\gg 1$
(\S\ref{iso_stable}, Eq. \ref{iso_stable_disp}). We  therefore 
seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with 
%we wish to seek a criterion independent of k 
%\begin{align}
%  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
%\end{align}
\begin{align}
  \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
  |\hat{\omega}\khat|=O(1). 
\end{align}
Then
\begin{align}
  \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
\end{align}
from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
that
\begin{align}
  \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
        q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
\end{align}
at marginal stability for perturbations with large $\khat$. 
For a thin disk $\epsilon\ll 1$, we infer 
\begin{align}\label{iso_vsi_cond}
  \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
  \beta_\mathrm{crit} 
\end{align}
is required for the fundamental VSI to operate. From our discussion
for isothermal perturbations (\S\ref{iso_discuss}), we expect
$\beta<\beta_\mathrm{crit}$ to be a sufficient condition for the
VSI. That is, when $\beta<\beta_\mathrm{crit}$, perturbations are
effectively isothermal. For a thin disk 
$\beta_\mathrm{crit}\ll 1$, which suggests that finite thermal
relaxation rapidly stabilizes the VSI.

### 2015-02-23 linear_iso.tex
new subsection growth time vs rad for mmsn

\subsubsection{Necessity of vertical shear for
  instability}\label{integral_relation} 
We can establish a necessary condition for instability by multiplying
Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
terms when integrating by parts, by assuming $W$ or
$W^\prime$ vanishes at the boundaries, or that the boundaries are 
sufficiently far away so that the boundary terms are negligible because of the
decaying background density with increasing height. Then,
\begin{align}
  &\hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat \notag\\
  &=\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
\end{align}
It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
shear. In Appendix \ref{global_vertiso} we reach a similar conclusion
in the global linear problem.

@@

\subsubsection{Stability in the absence of vertical shear}
When $q=0$, Eq. \ref{iso_ode3} is
\begin{align}\label{hermite_ode}
  \left[w(\hat{z})W^\prime \right]^\prime + nW
  w(\hat{z}) =0, 
\end{align}
where we have defined a new eigenvalue
\begin{align}
  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
\end{align} 
Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
that the kinetic energy density remain bounded at infinity, then  
\begin{align}
  W \propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to a real
eigenfrequency $\sigma$ such that
\begin{align}
  \left|\hat{\sigma}\right| = \sqrt{n}
  \left(1+\hat{k}^2\right)^{-1/2}. 
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
n-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height.

>>>>

\subsubsection{Necessity of vertical shear for
  instability}\label{integral_relation} 
We can establish a necessary condition for instability by multiplying
Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
terms when integrating by parts, by assuming $W$ or
$W^\prime$ vanishes at the boundaries, or that the boundaries are 
sufficiently far away so that the boundary terms are negligible because of the
decaying background density with increasing height. Then,
\begin{align}
  &\hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat \notag\\
  &=\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
\end{align}
It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
shear. A similar conclusion can be reached in the equivalent linear  
problem in global cylindrical disk geometry.

@@

\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
When $q=0$, Eq. \ref{iso_ode3} is
\begin{align}\label{hermite_ode}
  \left[w(\hat{z})W^\prime \right]^\prime + nW
  w(\hat{z}) =0, 
\end{align}
where we have defined a new eigenvalue
\begin{align}
  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
\end{align} 
Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
that the kinetic energy density remain bounded at infinity, then  
\begin{align}
  W \propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to a real
eigenfrequency $\sigma$ such that
\begin{align}
  \left|\hat{\sigma}\right| = \sqrt{n}
  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
n-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height.

We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
has also been qualitatively observed in numerical simulations performed by 
\cite{stoll14}.

### 2015-02-23 summary.tex
new subsection growth time vs rad for mmsn

We have neglected viscosity. \cite{nelson13} have shown that an
alpha viscosity of $\lesssim O(10^{-4})$ is required for the VSI to
operate. It is of interest to quantify the effect of viscosity
on the linear VSI. This would help identify, more accurately, the
appropriate regions in a protoplanetary disk that are sufficiently
laminar to develop the VSI.  This problem is, however, complicated by
the fact that proper inclusion of viscosity introduces viscous heating
in the energy equation.

>>>>

We have neglected viscosity. \cite{nelson13} have shown that an
alpha viscosity of $\lesssim O(10^{-4})$ is required for the VSI to
operate. It is of interest to quantify the effect of viscosity
on the linear VSI. This would help identify, more accurately, the
appropriate regions in a protoplanetary disk that are sufficiently
laminar to develop the VSI.  This problem is, however, complicated by
the fact that proper inclusion of viscosity introduces viscous heating
in the energy equation, as well as a meridional flow in the equilibrium state.
